---
title: 'Assignment: Due Monday, 11:59pm'
author: "Jared Wada"
date: "September 26, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
```

## 1. Brute force with only 5 keys

For this exercise, you will need to install the R package `digest` from CRAN. See the examples in the notes for how to perform AES encryption and decryption using this package.

Suppose that Eve knows that Alice has encrypted a plaintext block $m$ using AES in ECB mode. Suppose she also knows that Alice used one of the 5 128-bit keys given below, resulting in the ciphertext $c$.

```{r}
m <- charToRaw("myplaintextblock")
key1 <- charToRaw("collaborationist")
key2 <- charToRaw("eigenfrequencies")
key3 <- charToRaw("greatgrandmother")
key4 <- charToRaw("lighthousekeeper")
key5 <- charToRaw("multibillionaire")
keys <- list(key1, key2, key3, key4, key5)
c <- charToRaw("U\xa4\x90\020pt\037\xb2\xf8\xfaB`5\xb7$]")
```

Use the code block below to determine the key. Which key was used?

```{r}
library(digest)
## Show your attack here
for (i in 1:5){
key <- keys[[i]]
myAESobject <- AES(key, mode="ECB")
if(isTRUE(all.equal(myAESobject$encrypt(m),c))){print(rawToChar(key));
break;
}
}
```

## 2. Use `openssl` to make an ECB mode.

For this exercise, you will need to install the R package `openssl` from CRAN. This package has a function called `aes_cbc_encrypt`, but no function for ECB mode. Use the CBC mode (with an appropriate choice of initial vector) to write a function that encrypts a single 128-bit block in ECB mode. Test your function by comparing it to the ECB mode in the `digest` package.

```{r}

library(openssl)
## Show ECB function here and test it
aes_ecb_encrypt <- function(k, pt){
	return(aes_cbc_encrypt(data = pt, key = k, iv = raw(16))[1:(length(pt))])
}

m <- charToRaw("myplaintextblock")
key4 <- charToRaw("lighthousekeeper")

aO = AES(key = key4, mode = "ECB")
properECBText <- aO$encrypt(m)
myECBText <- aes_ecb_encrypt(k = key4, pt = m)

if(isTRUE(all.equal.raw(myECBText, properECBText))){
	print("Success!")
}else{
	print("Failure!")
}

## Show ECB function here and test it
```

## 3. Practice the Extended Euclidean Algorithm

Use the extended Euclidean algorithm to find the inverse of 7 in $\mathbb{Z}_{30}$. Show the steps of the Euclidean algorithm, show the backsubstitution, and check your answer by multiplying. 


$$30 = 7(4) + 2$$

$$7 = 2(3) + 1$$

$$2 = 1(2)$$

Therefore our GCD is 1.

Backsubstitution

$$1 = 7 - 3(2)$$

$$1 = 7 - 3(30 - 7(4))$$

$$30(-3) + (12 + 1)(7) = 1$$

$$1 = 30(-3) + 7(13)$$. 

In $\mathbb{Z}_{30}$, the -3 becomes 27

Check math. 

$$30(27) + 7(13) = 901$$ which leaves 1. 



## 4. Computing inverses in Galois Fields

Use the extended Euclidean algorithm to find the inverse of $x^2$ in the field $\mathbb{Z}_2[x]/(p(x))$, where $p(x) = x^4+x^3+x^2+x+1$. Show the steps of the Euclidean algorithm, show the backsubstitution, and check your answer by multiplying.

Euclidean algorithm

$${x^4+x^3+x^2+x+1} = x^2(x^2+x+1)+(x+1)$$

$x^2 = (x+1)(x-1) + 1$ The LCD of these two is 1

$$(x+1) = 1(x+1) + 0$$

Backsubstitution

$$1 = x^2 - (x+1)(x-1)$$

$$= x^2 - (x-1)(x^4+x^3+x^2+x+1) - (x^2(x^2+x+1))$$

In the ring $\mathbb{Z}_{2}$ $(x-1) = (x+1)$

$$= x^2 - {(x^4+x^3+x^2+x+1) - (x^2(x^2+x+1))}{(x^4+x^3+x^2+x+1) - (x^2(x^2+x+1))}$$

$$1 = ({x^4+x^3+x^2+x+1}{0}) + ({x^2}(1)$$

This is where I'm not sure where to proceed, but by filling out the inverse table, I found the answer to be $x^3$

## 5. Multiplication Cipher in ECB Mode

Define an encryption function on 4-bit blocks by representing 4-digit bit strings as polynomials in the field of Problem 3, and defining $E(p(x)) = x^2 p(x)$. Encrypt the 3-block string $001101010011$ using this cipher in ECB mode.

$$0011 = x+1$$ 

$$0101 = x^2+1$$

$$0011 = x+1$$

Then multiply each by $x^2$, convert back to binary and append.

$$x^2(x+1) =  x^3+x^2 = 1100$$

$$x^2(x^2+1) = x^3+x+1 = 1011$$

$$x^2(x+1) =  x^3+x^2 = 1100$$

The cipherstring is 110010111100

## 6. Multiplication Cipher in CBC Mode

Define an encryption function on 4-bit blocks by representing 4-digit bit strings as polynomials in the field of Problem 3, and defining $E(p(x)) = x^2 p(x)$. Encrypt the 3-block string $001101010011$ using this cipher in CBC mode, with an initial vector of $1001$. 


$$0011 xor 1001 = 1010$$ 
$$(x^3+x)x^2 = x^5+x^3 = x^3+1 = 1001$$
$$0101 xor 1001 = 1100$$
$$(x^3+x^2)x^2 = x^5+x^4 = x^3+x^2+x = 1110$$
$$0011 xor 1110 = 1101$$ 
$$(x^3+x^2+1)x^2 = x^3+x = 1010$$
The encrypted string is 100111101010
