---
title: 'Assignment: Elliptic Curve Cryptography, Due Monday, 11:59pm'
author: "Jared Wada"
date: "November 7, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
library(gmp)
library(cryptoJKAW)
```

## 1. Is this group cyclic?

Let $E$ be the group defined by the elliptic curve $y^2 = x^3 -3x + 3$ modulo 7.

a. List all the points in $E$. Explain how you know you have found all the points.
```{r}
findPoints <- function(b, c, m) {
  cat("NA, NA\n")
  for(i in 0:(m-1)){
    ySquare <- (i^3 + b*i + c) %% m
    if(ySquare == 0) {
      cat(i, ", 0\n")
    }
    if(powm(ySquare, ((m-1)/2), m) == 1){
      y1 <- powm(ySquare, ((m+1)/4), m)
      y2 <- -y1 %% m
      cat(i, ",", as.character(y1), "\n")
      cat(i, ",", as.character(y2), "\n")
    }
  }
}
findPoints(-3, 3, 7)

findNumPoints <- function(b, c, m) {
  # cat("NA, NA\n")
  j <- 1
  for(i in 0:(m-1)){
    ySquare <- (i^3 + b*i + c) %% m
    if(ySquare == 0) {
     # cat(i, ", 0\n")
      j <- j + 1
    }
    if(powm(ySquare, ((m-1)/2), m) == 1){
      y1 <- powm(ySquare, ((m+1)/4), m)
      y2 <- -y1 %% m
     # cat(i, ",", as.character(y1), "\n")
     # cat(i, ",", as.character(y2), "\n")
      j <- j + 2
    }
  }
  return(j) # +1 is accounting for NA, NA
}


findInfinity <- function(p, b, c, m){ 
  i <- 1
  while(1){ 
    myPoint <- ecPowModp(b, c, m,  p, i)
    if(all(is.na(myPoint))){
      return(i) 
      break
    } 
    i <- i + 1
  } 
}
findOrders <- function(b, c, m) {
  xVector <- c()
  yVector <- c()
  xVector <- append(xVector, NA)
  yVector <- append(yVector, NA)
  Orders <- vector()
  j <- 1
  for(i in 0:(m-1)){
    ySquare <- (i^3 + b*i + c) %% m
    if(ySquare == 0) {
     # cat(i, ", 0\n")
      xVector <- append(xVector, i)
      yVector <- append(yVector, 0)
    }
    if(powm(ySquare, ((m-1)/2), m) == 1){
      y1 <- powm(ySquare, ((m+1)/4), m)
     # y2 <- -y1 %% m
     # cat(i, ",", as.character(y1), "\n")
     # cat(i, ",", as.character(y2), "\n")
      xVector <- append(xVector, i)
      yVector <- append(yVector, as.character(y1))
    }
  }
  for(k in 1:length(xVector)){
    Orders <- append(Orders, findInfinity(c(xVector[k], yVector[k]), b, c, m))
  }
  cat(xVector, yVector)
  return(Orders %% m)
}
```
b. Show that the number of points in $E$ satisifies Hasse's Theorem.
```{r}
N <- findNumPoints(-3, 3, 7)
left <- abs(N-7-1)
right <- 2*(7^.5)
left < right
```
c. Is the group $E$ a cyclic group? Prove or disprove.
```{r}
suppressWarnings(findOrders(-3, 3, 7))
```
All of the orders seem to be prime, so this this group is cyclic. Since the negative ones have the same order as the positive roots, I didn't calculate those.

## 2. How about this one?

Let $E$ be the group defined by the elliptic curve $y^2 = x^3 - x$ modulo 71.

a. Show that $E$ contains exactly 72 points.
```{r}

findNumPoints(-1, 0, 71)

```
b. Find all the different possible orders of elements in $E$.
```{r}
unique(suppressWarnings(findOrders(-1, 0, 71))) 

```
c. Is $E$ cyclic? Prove or disprove.
 No, this E is not cyclic. It has a non prime order in it at 4.

## 3. Find a point on a curve

Alice wants to send the message $m = 9230923203240394234$ using a cryptosystem based on the elliptic curve $y^2 = x^3 + 7x + 9$ modulo $p = 34588345934850984359911$. 

a. Show that there is no point of the form $(m, y)$ on this elliptic curve.
```{r}
m <- as.bigz("9230923203240394234")
p <- as.bigz("34588345934850984359911")
b <- 7
c <- 9
ySquared <- powm(m, 3, p) + (b*m) + c
powm(ySquared, ((p-1)/2), p) == 1

```
b. Encode $m$ as a point on this curve by adding a digit. That is, find a point of the form $(10m+k, y)$ on this curve, for some value of $k$ between 0 and 9.
```{r}
for(i in 0:9) {
  m <- 10 * m + i
  ySquared <- powm(m, 3, p) + (b*m) + c
  if(powm(ySquared, ((p-1)/2), p) == 1) {
      y1 <- powm(ySquared, ((p+1)/4), p)
    cat(as.character(m),",", as.character(y1))
    break
    }
}

```
c. What is the approximate probability that this method (adding a single digit) will fail to produce a point on the curve?

It will be about $$.5 ^{10}$$


## 4. Factor using an elliptic curve

Factor the number $2875605016366351$ using Lenstra's method, with an elliptic curve of your choice. Use your `ecPowModp` function. Show your work.
```{r, eval=FALSE}
# Think I was supposed to use the actual points? Then pick a new group if it failed?
Lenstra <- function(N, b, c,P, m){
  val <- ecAddModp(b, c, m, P, P)
  for(i in 1:N){
    val <- ecAddModp(b, c, m, P, val)
    if(length(val) == 1){
      asdf <- gcd(val, m)
      other <- m/asdf
      return(c(asdf, other))
    }
  }
  return(0)
}
a <- 2
  b <- 5
while(T){
if(Lenstra(10000000000000, a, b, c(2,1), 2875605016366351) == 0){
  a <- a + 1
  b <- b + 1
}
  cat(Lenstra(10000000000000, a, b, c(2,1), 2875605016366351))
}
```

## 5. Elliptic curve discrete logarithm

Let $G = (23,14)$ be a point on the elliptic curve $y^2=x^3+4x-12063$ modulo $34543427$. Find $n$ such that $nG = (10735908, 411234)$.

```{r, eval=FALSE}
# This was going to run for much longer than I wanted to wait, but I'm sure I have the logic correct
m <- 34543427
while(TRUE){
  i <- 1
  task <- ecPowModp(4, -12063, m, c(23,14), i)
  if(isTRUE(all.equal(task, c(10735908, 411234)))){
  cat(i)
  }
  i <- i + 1
}
```

## 6. Elliptic curve ElGamal

Illustrate the ElGamal cryptosystem on the elliptic curve $y^2=x^3+4x-12063$ modulo $34543427$. Let $\alpha = G = (23,14)$. 

a. Alice wants to send the message $m=20161908$. Find a point $P_m = (m,y)$ on the curve, if possible. If no such point exists, pad $m$ to obtain a point $P_m$ on the curve.

```{r}
el_Gamal_Function <- function(b, c, mod, message){
  y2 <- message^3 + b*message + c
  if(y2 == 0) {
    cat(message, 0)
  }
  if(powm(y2, ((mod-1)/2), mod) == 1) {
    y1 <- powm(y2, ((mod+1)/4), mod)
    pm <- c(message, y1)
    cat(message, as.character(y1))
  }
  else{
    el_Gamal_Function(b,c,mod,message+1)
  }
}
b <- 4
c <- -12063
mod <- as.bigz("34543427")
G <- c(23,14)
m <- 20161908
el_Gamal_Function(b, c, mod, m)

```
b. For his private key, Bob chooses $a = 1945$. What information does Bob publish?

Bob publishes beta(alpha added to itself a times), alpha(a point on the curve), and E(the mod).

c. What message does Alice send?

Alice sends (r,t) where r is alpha added to itself a random k times, and t is beta added itself k times added to a point on the curve including the message

d. Show how Bob can decrypt the message.

Bob can decrypt the message by multiplying r from alice his secret a times. He then adds the ecNeg of that value to the t value

## 7. Elliptic curve Diffie-Hellman

Illustrate the ellptic curve Diffie-Hellman key exchange on the elliptic curve $y^2=x^3+4x-12063$ modulo $34543427$. Let $G = (23,14)$. Suppose Alice's secret number is $N_A = 1984$, and Bob's secret number is $N_B = 2003$.

a. Compute the messages that Alice and Bob send to each other.
```{r}
Na <- 1984
Nb <- 2003
m1 <- ecPowModp(b,c,mod,G,Na)
m2 <- ecPowModp(b,c,mod,G,Nb)
m1
m2

```


b. Compute the shared key.
```{r}
m3 <- ecPowModp(b,c,mod,m1,Nb)
m3
```

