---
title: 'Assignment: Due Monday, 11:59pm'
author: "Jared Wada"
date: "October 24, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
library(gmp)
library(openssl)
library(cryptoJKAW)
```

## 1. Birthday Theorem examples

1. Assuming that Social Security numbers behave as a random oracle, what is the probability that at least two people in a group of 350 have the same last four digits of their SSN's? Compute the answer exactly and approximately, using the formulas given in the Birthday Theorem.
```{r}
possibilities <- 10000
total <- 10000 - 349
p <- as.numeric("1")
for(i in total:9999){
  p <- p * as.numeric(i/possibilities)
}
ans <- (1 - p)
ans

r <- 350
n <- 20000
e <- exp(1)

a <- 1 - e^((-(r^2))/n)
a


```
2. Approximately how many messages would have to be hashed to do a birthday attack on SHA-256 that has a probability of 0.25 of succeeding? (Use the Birthday Theorem.)


```{r}
library(gmp)

x <- log(.75)
w <- log(1)
y <- pow.bigz(as.bigz("2"), as.bigz("257"))
p <- as.numeric(w-x) * as.numeric(y)
z <- p^.5
cat(z)

```





3. Approximately how many messages would have to be hashed to do a birthday attack on SHA-256 that has a probability of 0.99 of succeeding? (Use the Birthday Theorem.)

```{r}
library(gmp)

x <- log(.01)
p <- as.numeric(w-x) * as.numeric(y)
z <- p^.5
cat(z)

```






## 2. Find a collision 

Find a collision for your hash function `miniSHA`. That is, find two strings `x1` and `x2` such that `miniSHA(x1)` equals `miniSHA(x1)`. Use a birthday attack with randomness (everyone should get different answers). Do your work in a separate R script, outside of this file and your project. Once you find the collision, show the code that you used for the attack below, along with the strings that you found.

```{r, eval=FALSE}
# show your code here, but don't execute it here
library(random)
i <- 22
stringa <- "abc123_townshend 20"

collision <- function(){
  while(1) {
    stringb <- "abc123_townshend "
    
    stringb <- paste(stringb, i)
    #cat(stringa)
    #cat(stringb)
    if(miniSHA(stringa) == miniSHA(stringb)) {
      cat(sprintf("%s and %s are a match\n", stringa, stringb))
      cat(sprintf("%s", miniSHA(stringa)))
      break
    }
    i <- i + 1
  }
}
collision()


# Produces abc123_townshend 20 and abc123_townshend  151872 are a match
# they matching hash is26b4
```


## 3. Find a document with the same hash

Suppose Alice and Bob are using a signature scheme using the SHA-1 hash function, but instead of using the whole hash, they are just using the first two bytes of the SHA-1 hash. Bob sends the following message to Alice to sign.
```{r}
m1 <- "On the first day of 2017, I agree to pay Bob five dollars."
```

Explain how Bob could modify the following message `m2` to have the same two-byte hash as `m1` by adding no more than 16 additional whitespace characters between the words. Your answer should explain why Bob is guaranteed to find a collision. (Apply the Pigeonhole Principle.)

```{r}
m2 <- "On the first day of 2017, I agree to pay Bob one million dollars in small unmarked bills."
```


To get the similar two byte hash, he has to find a random hash for those two bytes, which could be any combination two 16 bit strings.
So the total amount of things he has to collide with is $16^2$ or $256$.
To modify the string by only adding whitespace between words with 16 whitespaces we do the calculation c(32, 16). 
This number is much, much greater than 256, just by looking, and so there should definitely be more than one "pigeon" per hole.
Thus a collision is guarenteed to be found.

## 4. Find a multicollision

Let $p = 95394590839549803540983759$. Consider the hash function $h(x)=x^2 \bmod p$, defined on the set of natural numbers. Find five different messages (i.e., numbers) that have the same hash. (Multiple collisions are called *multicollisions*.)
```{r}
p <- as.bigz("95394590839549803540983759")
additive <- 777
powm((p+additive), 2, p)
powm((2*p + additive), 2, p)
powm((3*p+additive), 2, p)
powm((4*p+additive), 2, p)
powm((5*p + additive), 2, p)
```

This works beacuse the hash function is so simple. You can just add the modulus to a number and then muliply that mod a few times to get the same exact hash. 


## 5. Authenticated Encryption

Suppose Alice and Bob both have public encryption functions $E_A$ and $E_B$, and private decryption functions $E_A^{-1}$ and $E_B^{-1}$. Let $m$ be a message that Alice wants to send to Bob. Alice sends $c = E_B(E_A^{-1}(m))$. Then Bob applies $E_A(E_B^{-1}(c))$.

1. Explain why Alice has access to the functions she used and Bob has access to the functions he used.

Alice has Bob's public $E_B$ and her own $E_A^{-1}$, so she can obviously send c, while providing her own m.
Bob can do his application because he has access to Alice's public $E_A$ and his own private $E_B^{-1}$.
This is how public encrytion functions work.

2. What will be the result of Bob's calculation?

Bob will decrypt c into the original message m.

3. Could anyone other than Alice have sent the message? How does Bob know?

No, because based on how public encryption works, only Alice should have the private key to encrypt the message properly, if Bob decrypted a coherent message.

4. Could anyone other than Bob have received (and decrypted) this message?

No, because they would also need Bob's private key. They could intercept it, but without Bob's key, it would be useless.

5. Make this algorithm into an authenticated encryption scheme using RSA. Since Alice and Bob both need private and public keys, start by making both Alice and Bob do steps 1 through 4 of the RSA algorithm (p. 165) to get $n_A, n_B$, etc. Then explain what Alice sends and what Bob does.

Alice and Bob both pick big primes p and q, and keep them as part of their own secret keys. They then calculate their own public n values.
Next they both pick an exponent e, also a public value. They then calculate their own private d values equal to the inverse of e in p-1. Since both peoples' n and e values are now public they can now do the authentification. Alice and Bob can each encrypt a message using each other's public n and e values and ask the other person to decrypt using their respective private d, p, and q values and verify that the message was correct.

