---
title: 'Assignment: Due Monday, 11:59pm'
author: "Jared Wada"
date: "September 19, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
```

## 1. Invertibility over $\mathbb{Z}_2$

Let $M$ be a $3\times 3$ matrix over $\mathbb{Z}_2$. Consider the following matrix.
$$
T = \begin{bmatrix}
1 & 0 & 1 & 0 & 1 & 0 & 1 \\
0 & 1 & 1 & 0 & 0 & 1 & 1 \\
0 & 0 & 0 & 1 & 1 & 1 & 1 
\end{bmatrix}
$$
Explain how and why the matrix $MT$ can be inspected to determine whether $M$ is invertible. If $M$ were an $n\times n$ matrix over $\mathbb{Z}_2$, what would the dimensions of $T$ have to be (in terms of $n$)?

If any column of MT has all zeroes, M is univertible. This is because by definition of an invertible matrix, if any vector is multiplied by a number other than zero, it is uninvertible. The dimensions of T would have to be $2^{n-1}\times n$. 

## 2. Decrypt the Hill Cipher

The following ciphertext was produced by a Hill cipher.
```
jghbqlxdvdohooikfcvqtkxljasmusjh
```
The key used was
$$
\begin{bmatrix}
16 & 6 & 1 & 3 \\
5 & 3 & 0 & 3 \\
10 & 3 & 1 & 0 \\
1 & 0 & 0 & 3
\end{bmatrix}
$$
Find the original plaintext. Give the matrix you used to perform the decryption. You can either do this using LaTeX, or by including a code block that prints the matrix.

The original plaintext is "whatisthecubsmagicnumberrightnow". I used this matrix
$$
\begin{bmatrix}
1 & 25 & 25 & 0 \\
16 & 19 & 10 & 17 \\
20 & 5 & 7 & 1 \\
17 & 9 & 9 & 9
\end{bmatrix}
$$
Also the magic number is 0. They already clinched.

## 3. Cracking an LFSR Sequence

The following bit string was produced by a linear recurrence.
$$
1 0 1 0 0 1 1 1 0 1 0
$$

1. List the shift matrices of dimensions $2\times 2$ to $6\times 6$. (The first is done for you, and the second is started, so you can see the syntax.)
     $$
     \begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}, 
     \begin{bmatrix} 1 & 0 & 1 \\ 0 & 1 & 0 \\ 1 & 0 & 0 \end{bmatrix},\begin{bmatrix} 1 & 0 & 1 & 0 \\ 0 & 1 & 0 & 0 \\ 1 & 0 & 0 & 1 \\ 0 & 0 & 1 & 1 \end{bmatrix},  \begin{bmatrix} 1 & 0 & 1 & 0 & 0 \\ 0 & 1 & 0 & 0 & 1 \\ 1 & 0 & 0 & 1 & 1 \\ 0 & 0 & 1 & 1 & 1 \\ 0 & 1 & 1 & 1 & 0 \end{bmatrix}, \begin{bmatrix} 1 & 0 & 1 & 0 & 0 & 1 \\ 0 & 1 & 0 & 0 & 1 & 1 \\ 1 & 0 & 0 & 1 & 1 & 1 \\ 0 & 0 & 1 & 1 & 1 & 0 \\ 0 & 1 & 1 & 1 & 0 & 1 \\ 1 & 1 & 1 & 0 & 1 & 0 \end{bmatrix}
     $$
2. Determine which of your shift matrices are invertible, then use the weakness theorem to make a conjecture about the length of the linear recurrence.

  The first two matrices are invertible, and the third is not. Going by the weakness theorem that means that the rest of the matrices are also uninvertible. This means that the linear recurrance is most likely 3. 
  
3. Find the linear recurrence that produced the sequence. You don't need to show all your work, but please give the system of equations that you solved and the solution to the system.

Using the second matrix for coefficient values. 
1x + 0y + 1z = 0. => x + z = 0
0x + 1y + 0z = 0. => y = 0
1x + 0y + 0z = 0. => x = 0
Using the first equation, z = 0.


## 4. Make a Good LFSR

In the code block below, write some code that produces a sequence of bits using a linear recurrence of length 7, such that the period of this sequence is greater than 100. Explain how you can tell that the period is greater than 100.
```{r}
# put your code here
x <- numeric(108)
x[1:7] <- c(1,0,0,1,1,1,0)
for(i in 1:101) {
  x[i+7] <- (x[i] + x[i+3]) %% 2 
}
firstSev <- x[1:7]
for(i in 8:108){
  samp <- x[i:(i+6)]
  print(samp)
  print(firstSev)
  if(isTRUE(all.equal(samp, firstSev))){
    print("There's a match")
    break
  }
  else print("None")
}
```
If this code worked, it should print "None" a bunch of times and show both the vectors to signal that it checked each subset of the 108 to make sure the first 7 digits didn't repeat. If not, it should tell me there was a copy by printing "there's a match" and stop. Since it checked over 100 variances, and had no repeats, the period should be over 100. 

