---
title: 'Assignment: Due Monday, 11:59pm'
author: "Sam McCollum"
date: "September 19, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
```

## 1. Invertibility over $\mathbb{Z}_2$

Let $M$ be a $3\times 3$ matrix over $\mathbb{Z}_2$. Consider the following matrix.
$$
T = \begin{bmatrix}
1 & 0 & 1 & 0 & 1 & 0 & 1 \\
0 & 1 & 1 & 0 & 0 & 1 & 1 \\
0 & 0 & 0 & 1 & 1 & 1 & 1 
\end{bmatrix}
$$
Explain how and why the matrix $MT$ can be inspected to determine whether $M$ is invertible. If $M$ were an $n\times n$ matrix over $\mathbb{Z}_2$, what would the dimensions of $T$ have to be (in terms of $n$)?

```
The resulting matrix will be invertible only if MT contains no zero columns
```
$$
sizeof(T) = n\times 2^{(n-1)}
$$

## 2. Decrypt the Hill Cipher

The following ciphertext was produced by a Hill cipher.
```
jghbqlxdvdohooikfcvqtkxljasmusjh
```
The key used was
$$
\begin{bmatrix}
16 & 6 & 1 & 3 \\
5 & 3 & 0 & 3 \\
10 & 3 & 1 & 0 \\
1 & 0 & 0 & 3
\end{bmatrix}
$$
Find the original plaintext. Give the matrix you used to perform the decryption. You can either do this using LaTeX, or by including a code block that prints the matrix.

```{r}
stringToMod26 <- function(x) {utf8ToInt(x)-utf8ToInt("a")}
mod26ToString <- function(x) {intToUtf8(x+utf8ToInt("a"))}
hillCipher <- function(txt, keyMatrix) {
  pt <- stringToMod26(txt)
  n <- attributes(keyMatrix)$dim[2]
  suppressWarnings(
    mPtxt <- matrix(pt,nrow=n) # repeats text so length is a multiple of n
  )
  mCtxt <- (keyMatrix %*% mPtxt) %% 26
  return(mod26ToString(as.vector(mCtxt)))
}
keyMatrix <- matrix(data = c(1, 25, 25, 0,
				16, 19, 10, 17,
				20, 5, 7, 1,
				17, 9, 9, 9), nrow = 4, ncol = 4, byrow = TRUE)
print(keyMatrix)
print(hillCipher("jghbqlxdvdohooikfcvqtkxljasmusjh", keyMatrix))
```

## 3. Cracking an LFSR Sequence

The following bit string was produced by a linear recurrence.
$$
1 0 1 0 0 1 1 1 0 1 0
$$

1. List the shift matrices of dimensions $2\times 2$ to $6\times 6$. (The first is done for you, and the second is started, so you can see the syntax.)
     $$
	\begin{bmatrix} 1 & 0 \\ 0 & 1 \end{bmatrix}, 
	\begin{bmatrix} 1 & 0 & 1 \\ 0 & 1 & 0 \\ 1 & 0 & 0 \end{bmatrix},
	\begin{bmatrix} 1 & 0 & 1 & 0 \\ 0 & 1 & 0 & 0 \\ 1 & 0 & 0 & 1 \\ 0 & 0 & 1 & 1 \end{bmatrix},
	\begin{bmatrix} 1 & 0 & 1 & 0 & 0 \\ 0 & 1 & 0 & 0 & 1 \\ 1 & 0 & 0 & 1 & 1 \\ 0 & 0 & 1 & 1 & 1 \\ 0 & 1 & 1 & 1 & 0 \end{bmatrix},
	\begin{bmatrix} 1 & 0 & 1 & 0 & 0 & 1 \\ 0 & 1 & 0 & 0 & 1 & 1 \\ 1 & 0 & 0 & 1 & 1 & 1 \\ 0 & 0 & 1 & 1 & 1 & 0 \\ 0 & 1 & 1 & 1 & 0 & 1 \\ 1 & 1 & 1 & 0 & 1 & 0 \end{bmatrix}
     $$
2. Determine which of your shift matrices are invertible, then use the weakness theorem to make a conjecture about the length of the linear recurrence.
```
	We think the length is 3 because the 4x4 matrix is the first one that is invertible
```
3. Find the linear recurrence that produced the sequence. You don't need to show all your work, but please give the system of equations that you solved and the solution to the system.
$$
	1a + 0b + 1c = 0\\
	0a + 1b + 0c = 0\\
	1a + 0b + 0c = 1\\
	0a + 0b + 1c = 1\\
	\\
	a = 1\\
	b = 0\\
	c = 1\\
$$

## 4. Make a Good LFSR

In the code block below, write some code that produces a sequence of bits using a linear recurrence of length 7, such that the period of this sequence is greater than 100. Explain how you can tell that the period is greater than 100.
```{r}
# put your code here
x <- numeric(200)
x[1:7] <- c(1, 0, 0, 1, 1, 1, 0)
for(i in 1:193) {
  x[i+7] <- (x[i] + x[i+3]) %% 2 
}
print(x)

# This function demonstrates that the first matching sequence occurs above the 100 character mark
for(i in 2:200){
	if( isTRUE(all.equal(x[1:7], x[i:(i+6)])) ){
		print(sprintf("Match: %d",i))
	}
}
```


