---
title: 'Assignment: Linear Codes, Due Wednesday 11/30, 11:59pm'
author: "Jared Wada"
date: "November 30, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
library(gmp)
library(cryptoJKAW)
```

## 1. Decode using a coset leader syndrome table

The following matrix is a generator matrix for a linear binary code.

```{r}
bch15 <- matrix(c(1,0,0,0,0,1,1,1,0,1,1,0,0,1,0,
                  0,1,0,0,0,0,1,1,1,0,1,1,0,0,1,
                  0,0,1,0,0,1,1,0,1,0,1,1,1,1,0,
                  0,0,0,1,0,0,1,1,0,1,0,1,1,1,1,
                  0,0,0,0,1,1,1,0,1,1,0,0,1,0,1),
                byrow=TRUE, nrow=5)
```

Answer the following questions, showing the mathematics or calculations that you used to determine your answers.

a. How many codewords are there in this code?
```{r}
2^(nrow(bch15))
```

b. How many errors can this code correct?
```{r}
(codeDistance(bch15)-1)/2
```
c. How many different syndromes are there?
```{r}
length(cosetLeaderSyndromeTable(bch15))
```
d. Bob receives the following message. Is it a codeword? If not, decode it to the nearest codeword. How many errors were corrected?

```{r}
mess <- c(1,0,1,1,0,1,0,0,0,1,0,0,1,1,1)
clst <- cosetLeaderSyndromeTable(bch15)
s <- syndrome(mess, parityCheckMatrix(bch15))
s %in% names(clst)
n <- match(s, names(clst))
slook <- clst[[n]]
as.numeric(xor(slook, mess))
hammingDistance((as.numeric(xor(slook, mess))), mess) # The number of errors corrected

```

## 2. Decode the Hamming $[63,57]$ code

The following commands compute the parity check matrix for the Hamming $[63,57]$ code. 

```{r}
m <- 6
powersOf2 <- 2^(0:(m-1))
ham63H <- sapply(c(setdiff(1:(2^m-1), powersOf2), powersOf2), 
                 function(x){as.numeric(intToBits(x)[1:m])})

```

Decode the following message (without computing the entire coset leader syndrome table). In which position was the error? Show your work.

```{r}
mess <- c(1,1,1,0,0,0,0,0,1,1,0,0,1,0,1,0,1,1,1,0,0,1,0,1,0,0,0,0,0,1,0,1,
          1,1,0,1,0,1,1,1,0,1,1,0,1,0,1,0,0,0,1,1,0,1,0,0,1,1,1,0,1,0,0)
syn <- syndrome(mess, ham63H)
syn <- as.numeric(unlist(strsplit(syn, "")))
post <- t(ham63H)
pos <- 0
for(i in 1:63){
  if(isTRUE(all.equal(syn, post[i,]))){
    pos <- i
    cat("The error position is:", i)
  }
}
mess[pos] <- mess[pos] + 1 %% 2
mess
```

## 3. Prove the lemma

Use the fact that $vH^T = \mathbf{0}$ for any codeword $v$ to prove both parts of the following lemma.

> **Lemma.** Let $C$ be a sytematic linear binary code, and let $H$ be its parity check matrix.
>
> 1. If $C$ contains a codeword of weight 1, then $H$ contains a zero column.
> 2. If $C$ contains a codeword of weight 2, then $H$ has two identical columns.

## 4. Build a Golay decoder

The Golay code $G_{24}$ has the following generator matrix.

```{r}
golay24 <- matrix(c(1,0,0,0,0,0,0,0,0,0,0,0,1,1,1,0,1,1,1,0,0,0,1,0,
                    0,1,0,0,0,0,0,0,0,0,0,0,1,0,1,1,0,1,1,1,0,0,0,1,
                    0,0,1,0,0,0,0,0,0,0,0,0,1,1,0,1,1,0,1,1,1,0,0,0,
                    0,0,0,1,0,0,0,0,0,0,0,0,1,0,1,0,1,1,0,1,1,1,0,0,
                    0,0,0,0,1,0,0,0,0,0,0,0,1,0,0,1,0,1,1,0,1,1,1,0,
                    0,0,0,0,0,1,0,0,0,0,0,0,1,0,0,0,1,0,1,1,0,1,1,1,
                    0,0,0,0,0,0,1,0,0,0,0,0,1,1,0,0,0,1,0,1,1,0,1,1,
                    0,0,0,0,0,0,0,1,0,0,0,0,1,1,1,0,0,0,1,0,1,1,0,1,
                    0,0,0,0,0,0,0,0,1,0,0,0,1,1,1,1,0,0,0,1,0,1,1,0,
                    0,0,0,0,0,0,0,0,0,1,0,0,1,0,1,1,1,0,0,0,1,0,1,1,
                    0,0,0,0,0,0,0,0,0,0,1,0,1,1,0,1,1,1,0,0,0,1,0,1,
                    0,0,0,0,0,0,0,0,0,0,0,1,0,1,1,1,1,1,1,1,1,1,1,1),
                  byrow=TRUE, nrow=12)
```

a. Use the fact that this code is a linear code to compute its minimum distance. How many errors can this code correct? (Show your work. You probably want to avoid using your `codeDistance` function.)

```{r}
rowSums(golay24)
# The minimum distance is 8, because it is the least count of ones among the rows.
# The # errors it can correct is 3 from (8 - 1) /2
```


b. Use the following code block to compute the coset leader syndrome table for this code. It may take a few minutes, so the following code block is set not to evaluate when this file is knitted. The second line of this code block saves the result to a file so you don't have to do this computation again, in case you need it for a future assignment. Change the path and name of the file `~/Downloads/filename.RData` to something appropriate. How much memory does this table occupy in the R environment? (Look in the environment tab.) How big is the stored file?

```{r, eval=FALSE}
golayTable <- cosetLeaderSyndromeTable(golay24)
save(golayTable, file="~/Downloads/filename.RData")
```

c. Decode the following message. How many errors were corrected?

```{r}
mess <- c(1,1,1,0,0,0,1,0,1,0,0,0,0,1,0,1,0,0,0,0,1,0,1,1)
sy <- syndrome(mess, t(golay24))
sy
B <- golay24[1:nrow(golay24), 13:ncol(golay24)]
sy2 <- syndrome(as.numeric(unlist(strsplit(sy, ""))), B)
e <- c(0,0,0,0,0,0,0,0,0,0,0,0)
e <- append(e,sy2)
e <- as.numeric(unlist(strsplit(e, "")))
ans <- (e + mess) %% 2
ans
hammingDistance(ans, mess)
```




