---
title: 'Assignment: Due Wednesday, 11:59pm'
author: "Jared Wada"
date: "September 28, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
library(gmp)
```

## 1. Multiplication Cipher in CTR Mode

Define an encryption function on 4-bit blocks by representing 4-digit bit strings as polynomials in the field of Problem 3, and defining $E(p(x)) = x^2 p(x)$. Encrypt the 3-block string $001101010011$ using this cipher in CTR mode, with an initial vector of $1001$. (For the purposes of this problem, let $L_8$ be the identity function.) Show all work. (Warning: The counter in CTR mode uses binary addition, while the $\oplus$ operation corresponds to addition in the field. These are different operations. For example, in this problem, $X_1 = 1001$, so $X_2 = 1001+1=1010$. As elements of the field, these bit strings correspond to $X_1 = x^3+1$ and $X_2 = x^3+x$.)

Three byte blocks are 0011, 0101, 0011. 

The encryption function is $E(p(x)) = x^2 p(x)$

The initial vector is 1001.

\begin{center}Block 1 \end{center}

$$X_1 \times x^2$$

$$1001\times x^2 = x+1\times x^2 = x^5+x^2 = 1 + x^2 = 0101$$

$$C_1 = X_1 \oplus B_1$$

$$0101 \oplus 0011 = 0110$$

$$C_1 = 0110$$

$$1001 + 1 = 1010$$

$$X_2 = 1010$$

\begin{center}Block 2\end{center}

$$X_2 \times x^2$$

$$1010\times x^2 = (x^3+x)\times x^2 = x^5+x^3 = x^3 + 1 = 1001$$

$$C_2 = X_2 \oplus B_2$$

$$1001 \oplus 0101 = 1100$$

$$C_2 = 1100$$

$$1010 + 1 = 1011$$

$$X_3 = 1011$$

\begin{center}Block 3 \end{center}

$$X_3 \times x^2$$

$$1011\times x^2 = (x^3+x+1)\times x^2 = x^5+x^3+x^2 = x^3+x^2+1 = 1101$$

$$C_3 = X_3 \oplus B_3$$

$$1101 \oplus 0011 = 1110 $$

$$C_3 = 1110$$

Then encryption string is $011011001110$.





## 2. Encrypt using RSA

In the code block below, encrypt the integer message using the RSA algorithm, with the given modulus and exponent.

```{r}

#' Convert string to big integer
#'
#' Converts a string to a bigz integer. First the characters of the string are converted to raw, then the
#' raw (hexadecimal) vector is converted to an integer, where the place values of this vector are assigned
#' from right to left.
#'
#' @param txt A character string
#'
#' @return A `bigz` integer
#' @export
#'
#' @examples
#' stringToBigz("A string, which may contain punctuation.")
stringToBigz <- function(txt) {
  nraw <- charToRaw(txt)
  l <- length(nraw)
  return(sum(as.bigz(256)^(l-(1:l))*as.numeric(nraw)))
}

#' Convert (some) big integers to strings
#'
#' This function is intended to serve as an inverse to the \code{\link{stringToBigz}} function.
#' If the raw representation of `n` contains 00's, this function will produce an
#' `embedded nul in string` error. Thus it is not suitable for all integers.
#'
#' @param n A `bigz` integer, whose raw representation must not contain 00's
#'
#' @return A character string
#' @export
#'
#' @examples
#' stringAsInt <- stringToBigz("Any string.")
#' bigzToString(stringAsInt)
#' # bigzToString(256^3+256) # embedded nul error!
bigzToString <- function(n) {
  numbytes <- ceiling(log2.bigz(n)/8)
  nnumeric <- numeric(numbytes)
  for(i in 0:(numbytes-1)) {
    b <- as.numeric(mod.bigz(n, 256))
    n <- divq.bigz(n, 256)
    nnumeric[numbytes-i] <- b
  }
  return(rawToChar(as.raw(nnumeric)))
}
library(gmp)
rsaModulus <- as.bigz("195788142788358164233804492957398488634726254
80010690459681625329893085627204988093818307212739677049956352567177
87746099255774603493751657981591279103676377321834690778523173801439
56666453740761405761559112081931710294389007965357426686841571207915
63250256451028748849392888561065149545396193209901120150922222695733
85885218236653180615298633266570810577713313325153177729917576091579
04367035499790847594316987574068066165228272596810445659823407499566
73617955415896513151098470897855653135143572581766041083134024184173
37290933287861015998551958245267490476692680563890349711670121990426
2326557198534825303109914717")
rsaExponent <- as.bigz("65537")
rsaMessage <- as.bigz("453787497332492283997614676595994471")

gcd.bigz(rsaExponent, rsaModulus) #Just checking the gcd is 1

encryption <- mod.bigz(pow.bigz(rsaMessage, rsaExponent), rsaModulus)
print(encryption)
```

Notice that, for this example, the ciphertext number has many more digits than the plaintext number. Will the RSA ciphertext always be longer than the plaintext? Explain.

The ciphertext will only have more digits if the chosen modulus has more digits than the plaintext.
This should be obvious, and the modulus will circle the number within itself. 
However, the modulus should be the product of two very large prime numbers, so unless the plaintext is very long, 
the ciphertext should be longer than the plaintext. 

## 3. Decrypt using RSA

Suppose that the following ciphertext (number) was encrypted using RSA with the modulus and exponent from Problem 2. 

```{r}
rsaCiphertext <- as.bigz("16053542953097522034129794649814372515869362991732174439829
3776550795524368337578956746793355042030038205926539079142418490671465617605065354907
4846969694778601218848152236677611918504345262491414443232089922643358336946362484478
7840597302531115865062721545459882263488079496144532695399516588916034043077740431329
3679140154219046246386812621840178092960797765021778623868350200999052841894073701215
5233606208746477785068289264069257218137457317674139071451943085300805611973907560305
7844509912461437119592469825499225257447545934250164343715835969613074707612423475222
013041220244987344227734563376080328735796169508")
```

Decrypt. You will need to use the fact that the modulus is the product of the two primes given below. Show the required calculations in the code block below. Use the `bigzToString` function to recover the original ciphertext.

```{r}
prime1 <- as.bigz("144911861660469186676135294469223926436941190769646690446708204735
8736042802470485903802764162570319046524677116495722587569335044623035240248379011509
3185063138983627389612280537711144526516541647238439825051192285576632558346504222927
4390742421922958707226985795059825145780503339835934302622599390902434169")
prime2 <- as.bigz("135108431114557702161176938735198129586127410854075454085216219632
8196814775832631024605026656139919348031710417226631499441380055668631715847496632676
0910748676123466333816030590962886084101772115154314805240281171448508474907855212648
7326570961868801681549328707734385122713774204115903767411319701066873093")

#Storing this value as it is used in multiple steps
factor <- mul.bigz(sub.bigz(prime1, 1), sub.bigz(prime2, 1)) 

#Finding the inverse of e
gcdex(rsaExponent,mul.bigz(sub.bigz(prime1, 1), sub.bigz(prime2, 1)))

#Finding the value of d, the inverse of e mod (prime1-1*prime2-1)
d <-mod.bigz(gcdex(rsaExponent,factor)[2], factor)

#Checking to see if d*e is 1 mod N
 mod.bigz(mul.bigz(d, rsaExponent), factor)
 
 #Running the decryption
 plaintext <- bigzToString(powm(rsaCiphertext, d, rsaModulus))
 
 plaintext

```

