---
title: 'Assignment: Due Monday, 11:59pm'
author: "Jared Wada"
date: "October 2, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(comment = NA)
library(gmp)
```

## 1. Factoring $pq$ when $p$ and $q$ are nearby primes.

Consider the following (composite) number.

```{r}
library(cryptoJKAW)
bigComp <- as.bigz("41495155688809929585124078636911611510124462322424
                   368999956573296906528114129081463997081530206029675
                   796566441088639070653584409712811422453347196222375
                   13761725653531896260925123613")
FermatTest(bigComp, as.bigz("234234"))
log10.bigz(bigComp) #number of digits (remove decimal numbers)
log2.bigz(bigComp) #number of bits (remove decimal numbers)
```

a. In the codeblock above, check that this number is composite by using your Fermat test.
b. In the codeblock above, find the number of digits in this number and the number of bits using logs.
c. Notice that the square of the following number is close to `bigComp`.  Use this information, and the fact that the factors of `bigComp` are approximately the same size, to factor `bigComp`. Show your code in the codeblock below.
   
```{r}
sqRoot <- as.bigz("203703597633448608626844568840937816105146839366593
                  6250636140449354381299763336706183397647")
stwo <- mul.bigz(sqRoot, sqRoot)
sub.bigz(stwo, bigComp) #square root of this is 114
n <- mul.bigz(sub.bigz(sqRoot, 114), add.bigz(sqRoot, 114))
n == bigComp
```

## 2. Find a pseudoprime that is not strong

Use your Fermat primality test to find all of the pseudoprimes between 1000 and 10000 for the base 7. Use the `isprime` function to determine that your pseudoprimes are composite.

```{r}
# Put your code for this problem here
for(i in 1000:10000){
  if(FermatTest(i, 7)){
    if(isprime(i) == F){
      cat(sprintf("%d is a pseudoprime\n", i))
    }
  }
}
```


## 3. Quick power calculation

Let $a$ be an integer. Suppose you have a calculator that can only perform the operations of squaring and multiplying. Show (using algebra) that you can calculate $a^{257}$ using only five calculator operations.

$$ a\times a = a^2$$
$$ (a^2)^2 = a^4$$
$$ (a^4)^2 = a^8$$
$$ (a^8)^2 = a^{16}$$
$$ (a^{16})^2 = a^{32}$$
$$ (a^{32})^2 = a^{64}$$
$$ (a^{64})^2 = a^{128}$$
$$ (a^{128})^2 = a^{256}$$
$$ a^{256} * a = a^{257}$$

## 4. Timing the `factorize` function

The `gmp` library has a function called `factorize` which will attempt to factor a large number. In R, you can see how long a command takes by using the `system_time` function:

```{r}
system.time(factorize(as.bigz("74534598735433245597")))

system.time(factorize(as.bigz("221073920338411049151031")))
system.time(factorize(as.bigz("1326443525118841866260617")))
system.time(factorize(as.bigz("7958661137742004364881321")))
system.time(factorize(as.bigz("47751966754183036233218137")))
system.time(factorize(as.bigz("286511800375000558445327027")))
system.time(factorize(as.bigz("1719070802400095237189887909")))
system.time(factorize(as.bigz("10314424803143471989157992487")))
system.time(factorize(as.bigz("371319292755115219695668554559")))
```

Use this command to determine the time it takes to factor the following numbers, each of which is a product of large primes. Are these times increasing linearly with the number of digits? If not, how are these times increasing?

```
221073920338411049151031, 1326443525118841866260617, 
7958661137742004364881321, 47751966754183036233218137, 
286511800375000558445327027, 1719070802400095237189887909,
10314424803143471989157992487,
371319292755115219695668554559
```

\begin{center} No, it doesn't seem to be increasing linearly according to the number of digits. \n
I can't find a reasonable relationship between the numbers. It seemed to grow slightly, or shrink slighty, and then at the end grew exponentially\end{center}

## 5. Short plaintext attack

```{r}
rsaMod <- as.bigz("2270785812851607857752236173866598165780980999480087
                  96470995669001062758570837819275351571519690240251463
                  92877047548296429131296960565225703131554564424303895
                  83920197030931790925387823837277203838574301774350132
                  42444504076066258379017983988851994235746436107443245
                  62528276931827066473866615454887905670524424503707688
                  92681879411999393643108663016006507255489833947310000
                  87923251226854295095181659541311283989229588168697957
                  58667406044391829713985703446445013454864090920545268
                  59941485219860927547608296980025433657621871715162467
                  84071119823459209996599672653802424029963364148197234
                  56853027238543378738028207982298417")
rsaExp <- as.bigz("65537")
c <- as.bigz("128216972055960183378721787714469111306799737814971002198
             6229586389359344497658184213712429285822188549926995588183
             3237258784376795565986432565955032055406700435350884952772
             8327573356968001432140335916137748548203476259553293133733
             6984514346587698999980727268439793304759211907286227159763
             8991440842731415663378367587187983360075334687571117158156
             7251718603971567669618892639720277733050807370950305289557
             6032600301346639961168091352569266230799971665474413926527
             4217957686818400851397262197048091402851157186436667777674
             1004576470040519183572125063272357752504338431403925758867
             97119952853745267282047381070641950241")

x <- list()
y <- list()

# The book said to take 10^9 numbers and check them, which was taking forever to compile. 
# This code is likely similar to Chris'(and anyone he may have helped) as he pointed me towards the intersect method
# after many failed attempts with for loops and sapplys.

for(i in 1:(10**9)){
  x[[i]] <- (c*powm(i, -rsaExp, rsaMod))
  y[[i]] <- powm(i, rsaExp, rsaMod)
  if(length(intersect(x,y)) == 1){
    xpos <- match(intersect(x,y), x)[[1]]
    print(xpos)
    ypos <- match(intersect(x,y), y)[[1]]
    print(ypos)
    print((xpos*ypos)%%rsaMod)
    print(bigzToString((xpos*ypos) %% rsaMod))
    break
  }
}


```

In the code block above, conduct a short plaintext attack to find the plaintext that produced the ciphertext `c`. (Follow the procedure on page 172 of [T].)



